The Essence of Summarization\\ in Pushdown Flow Analyses\\ \textbf{Supplemental Material}

### 2014-03-26 content.tex
More related work, spell checking.

%%
This paper gives a common, simple methodology to derive CFA2, PDCFA and its garbage-collecting extension in a manner similar to AAM.
%
Our presentation is meant to be a tutorial to this related literature, with the additional promise of a novel analysis.
%
We don't presume a language of formal semantics such that our technique applies to all formalisms in such a language.
%
An overly formal approach is unlikely to capture all ``real language's'' formalizations.
%
The additional abstraction also detracts from the operational intuitions and simplicity of the methodology.
%
We still present the definitions and proofs in a manner that should be well-positioned for mechanization.
%
The reader unfamiliar with or uninterested in correctness proofs should still be able to grok the technique and apply it with little difficulty.
%
This paper has been partially mechanized in the Coq proof assistant, with proofs in the supplemental materials\footnote{\url{http://github.com/ianj/concrete-summaries} in \texttt{model.v}}.

@@

CFA2 is an online version of a technique called ``summarization'' from Sharir and Pnueli \citep[Chapter 7]{local:muchnick:jones:flow-analysis:1981}, which is synonymous with PDCFA's crucial ``$\epsilon$-closure graph''.
%
The online quality of the algorithm is what makes it suitable for higher-order languages, since CFA essentially interprets the program for each discovered state to build the model on-the-fly.
%
Summarization algorithms need not be restricted to languages with well-bracketed calls and returns.
%
The technique is adaptable for higher precision in the common case but to still handle difficult cases such as first-class control.
%
This was shown for the \rackett{call/cc} operator by~\citet{ianjohnson:Vardoulakis2011Pushdown}.
%
This work illuminated the fact that we can harness the enhanced technology of pushdown analyses in non-pushdown models of computation.
%
Doing this sacrifices call/return matching in the general case, but in practice the precision is much better than the alternative regular model that, \eg, 0-CFA would provide.
%%

@@

%%
The second piece of machinery serves two purposes:
%
(1) the state space contains all function callsites to project out a control-flow graph, and %
(2) the worklist semantics of execution makes it possible to appeal to an incomplete memo table, so the saved stacks allow new results to flow to calls that used incomplete memo information.
%
Once the machine semantics is in this form, simple point-wise abstraction leads to the summarization algorithms that we see in the literature.

@@

\paragraph{Notation:} we make judicious use of the isomorphism between associations of tuples to spare the paper from excessive parentheses, even though our implementations are in a Lisp-descendent language.

@@

%%
The tail of a continuation within a function is some $\krt{\mctx}$, which instructs the semantics to memoize the result of the function evaluation at $\mctx$.
%
A context, $\mctx$, must contain enough information for memoization to be sound, \ie, any extra components not in $\mctx$ must be irrelevant to the evaluation of the function.
%
In this case, the context includes the function, environment, and store.
%
All the extra components at a context $\mctx$ get stored in a separate table, $\mktab$, so that updates to the memo table will propagate to all callers that depend on the result of $\mctx$.
%%

%%
Continuations count as ``extra'' in this semantics since functions cannot inspect the stack, so $\krt{\mctx}$ might be considered a special allocation strategy for continuations in the AAM viewpoint.
%
We separate the table of continuations, $\mktab$, from the store, $\mstore$, since $\rt$ continuations contain stores---this leads to a recursive data-structure that we are trying to avoid.
%
The store is an important ingredient to maintaining enough precision to keep a pushdown abstraction.
%
The semantic spaces for the machine are modified as follows:

\begin{align*}
  \mstate \in \State &::= \tpl{\mexpr,\menv,\mstore,\mkont,\mktab,\mmemo} \\
  \mkont \in \Kont &::= \kmt \alt \krt{\mexpr,\menv,\mstore} \alt \kcons{\mkframe}{\mkont} \\
  \mktab \in \KTab &= Expr \times \Env\times \Store \to \wp(\Kont) \\
  \mmemo \in \Memo &= \Expr \times \Env\times \Store \to \wp(\Value \times \Store)
\end{align*}

\paragraph{The role of summaries:} notice the additional $\Memo$ component.
%
A ``summary edge'' in CFA2 or equivalently an $\epsilon$-edge in PDCFA is an edge from the source of a push edge to the target of the matching pop edge.
%
``Matching'' here means there is a path through machine reductions that don't change the stack, or through summary edges.
%
In our model, we only ``push'' when we call a function, and we only ``pop'' when we return with a value.
%
There is an analogy to something more operational: summary edges embody memoization.
%
Instead of following an entire path through a call to return a value, we simply jump from the call to the return with the result of the call, with the resulting context changes---in this case only the store (\todo{show in figure})
%%

@@

This semantics is only correct for fresh allocation strategies, which in turn make the reduction relation deterministic.
%
As it is, the CESK machine approximates the tabular machine, but not the other way around;
%
if we use a finite allocation strategy for example, then we don't explore some non-deterministic paths that lead to different results for a function call.
%
We thus explore all paths and share $\mktab$ and $\mmemo$ between all paths.

@@

\paragraph{Technical aside} There are minor differences with PDCFA's algorithm due to our choice to process one state at a time. If we process the entire frontier in one step, then we need to have additional tables of ``changes to $\mktab$ and $\mmemo$'' that need to be matched up in the following step. Here ``matched up'' means if both tables have the same context in their domain, we must add the state that has the memoized result(s) with the saved continuation(s). If we don't, then some memoized results may not flow to newly reached calling contexts. This corresponds with PDCFA's worksets of edges.
%

{
\setlength{\abovedisplayskip}{0pt}
\setlength{\belowdisplayskip}{4pt}
\setlength{\abovedisplayshortskip}{0pt}
\setlength{\belowdisplayshortskip}{8pt}
\begin{align*}
  \mastate \in \sa{State} &::= \tpl{\mexpr,\menv,\mstore,\mkont} \\
  \System &= \wp(\sa{State}) \times \wp(\sa{State} \times \sa{State}) \times \KTab \times \Memo \\
  {\mathcal F} &: \sa{\State} \to \System \to \System
\end{align*}
\begin{align*}
  {\mathcal F}(\mastate_0)(\emptyset, R, \mktab, \mmemo) &= (\emptyset, R, \mktab, \mmemo) \\
  {\mathcal F}(\mastate_0)(F\uplus\set{\mastate}, R, \mktab, \mmemo) &= (F\cup F', R \cup R', \mktab', \mmemo') \\
  \text{where } I &= \setbuild{\mstate'}{\wn(\mastate,\mktab,\mmemo) \stepto \mstate'} \\
                \wn(\tpl{\mexpr,\menv,\mstore,\mkont},\mktab,\mmemo)
                   &= \tpl{\mexpr,\menv,\mstore,\mkont,\mktab,\mmemo}
\end{align*}
\begin{align*}
    R' &= \setbuild{(\mastate,\mastate')}{\wn(\mastate', \_, \_) \in I \wedge (\mastate,\mastate') \text{ not a call using } \mmemo} \\
    \mktab' &=  \bigsqcup\setbuild{\mktab'}{\wn(\_, \mktab', \_) \in I} \\
    \mmemo' &= \bigsqcup\setbuild{\mmemo'}{\wn(\_, \_, \mmemo') \in I} \\
    F' &= (\setbuild{\mastate'}{\wn(\mastate', \_, \_) \in I} \cup \set{\mastate_0})\setminus (\pi_1 R \cup \pi_2 R)
  \end{align*}
}
%%

%%
A $\System$ embodies the reductions seen disregarding the administrative pop steps, $R$, the frontier set of states yet to analyze, $F$, the continuation table $\mktab$ and the memo table $\mmemo$.
%
The set of reductions does not include uses of the memo table since those do not correspond to true steps that the original machine takes.
%
We define $\pi_1$ and $\pi_2$ as the pointwise lifting of the left and right projections out of a pair, respectively.
% Note: Only if widened store.{
%
%The memo table need only map to values, since the global store will always subsume the store that would otherwise be mapped.
%
%All the states in the frontier are stepped with the current store, after which the next store is the least upper bound of all the resulting stores.
%
%The next frontier contains only states resulting from stepping the previous frontier that we haven't seen at this next store.}
%
The least fixed-point of ${\mathcal F}(\tpl{\mexpr,\bot,\bot,\kmt})$ is the analysis of closed expression $\mexpr$.
%
Note the ordering on systems is pointwise, so the bottom element is $(\emptyset,\emptyset,\bot,\bot)$, which we'll call $\mathit{sys}_0$.
%
Systematic techniques for a performant implementation can be found in \citet{ianjohnson:oaam:icfp2013}.
%

@@

\paragraph{Example:} Let's consider a monovariant allocation strategy to run the following example.
Note that monovariance means the address for each binding is the variable name itself, so $\menv$ is always the identity function. We therefore omit it.

@@

%%
We should expect that a pushdown analysis would predict this evaluates to true, and there are no loops in the program.
%
0-CFA~\citep{dvanhorn:Shivers:1991:CFA} claims there is a loop from the second call of \texttt{id} to the first, and thus predicts this program evaluates to true or false.
%
PDCFA claims there is no loop and that the result is true.
%
The example evaluation is in \autoref{fig:ex-eval}.
%
Here $\mkont_0 = \mathbf{lt}({\tt y}, {\tt (let\ ([z\ (id\ 1)])\ (}\le\ {\tt y}\ {\tt z))})$.
%
We maintain enough context to distinguish the return points of \texttt{id} to not rebind \texttt{y} to 1.
%
When determining control flow through the expression, we consult $\mktab$ to continue past function returns, so there is no confusion about a back edge from the second call to \texttt{id}.
%

@@

Garbage collection ($\Gamma$) restricts the heap to just the reachable addresses, so a semantics with GC introduces the following reduction rule:
\begin{center}
  \begin{minipage}{0.3\linewidth}
    \begin{align*}
      \mstate \stepto \Gamma(\mstate)
    \end{align*}
  \end{minipage}
  \begin{minipage}{0.65\linewidth}
    \begin{align*}
      \Gamma(\tpl{\mexpr,\menv,\mstore,\mkont, \mktab, \mmemo}) &= \tpl{\mexpr,\menv,\mstore|_{\reaches(\mathit{root}, \mstore)}, \mkont, \mktab, \mmemo} \\
      \text{where } \mathit{root} &= \touchesk{\mkont}(\mktab) \cup
      \mathit{range}(\menv|_{\mathit{fv}(\mexpr)})
    \end{align*}
  \end{minipage}
\end{center}
%
The ``reaches'' metafunction ($\reaches$) is a transitive closure of ``touches'' as it weaves through the store, starting from some root set of addresses:

@@

This semantics of course enjoys the soundness properties that GC does not introduce dangling pointers, equal length traces are equal up to GC (in the concrete), and abstractions are sound with respect to collected states.
%
The proofs are all similar to the original abstract GC paper's.
%
Technically since this semantics inspects the touched addresses of the continuation, the set of touched addresses should be considered part of the ``context'' in the $\mathbf{rt}$ continuations.
%
This extra context can be overlooked, but it leads to larger live address sets.
%
More live addresses means a less effective GC, but the introspective pushdown analysis chose this tradeoff for a possibly smaller state space.
%

@@

\begin{enumerate}
\item{stack allocation for some bindings in an additional $\mframe \in \Store$}
\item{strong updates on stack frames for looked-up addresses}
\end{enumerate}
There are two orthogonal features of the semantics, embodied by our semantics' two metafunctions ($\bind$ and $\lookup$, respectively in \autoref{fig:frame-semantics}).

@@

%%
The stack frame introduces a refinement over the original CESK machine, so we first prove it correct with respect to the concrete semantics, in a similar fashion to PDCFA.
%
\begin{theorem}[Correctness of stack frame refinement]\label{thm:refinement}
  The CES$\xi$K machine has a stuttering bisimulation refinement with the CESK machine with fresh allocation.
\end{theorem}
This follows from the invariant that if a stack frame maps $\maddr$ to $\mval$, then the store maps $\maddr$ to $\mval$, meaning that the special lookup function is equivalent to lookup directly in the store (exploits immutability).
%
We stutter past the stack frame replacement steps.
%
%That is, if we always allocated bindings in the heap.
%
%The additional escape information that allows us to elide allocation in the heap introduces a second level of reasoning, where the burden of proof falls on the escape analysis.
%
%The escape analysis adds information to use in the proof, that if $\mvar$ never escapes, then every reference to allocations of $\mvar$'s binding will always be able to use $\mframe$ for the lookup.
%
%With this proposition, we can prove that the CES$\xi$K machine that always allocates in the store has a bisimulation with the refined machine that exploits escape analysis to avoid unnecessary changes to the store.
%

@@

The two metafunctions, $\domark$ and $\domarkset$ are what do the heavy-lifting of adding marks to a continuation and reifying the mark set into a map of keys to lists.
%
Adding a mark to a continuation has an implicit equality check; if $\mval_k = \mval_k'$ concretely, then {\tt wcm} should do a strong update on $\mmarks$, but otherwise it should join since they could be two different values.
%
Concrete equality analysis is outside the scope of this paper, but doable (\eg, with abstract counting~\citep{dvanhorn:Might:2006:GammaCFA}), so we elide the details and call this strong-update-or-join operation $\sqcup_{!{}}$.
%
Marking a return continuation means we need to crawl through the table, which again can be circular, so we compute the least fixed point of ${\mathcal M}$ to get an extended table holding the marked continuations.
%
\newcommand{\rewritectx}{\mathit{rw}}
\newcommand{\domarkstep}{\mathit{mark}_0}
\begin{align*}
  \domark(\mkont, \mktab, \mval_k, \mval_v) &= (\mkont', \mktab) \text{ if } (\mkont',\bot) = \domarkstep(\mkont) \\
  \domark(\mkont, \mktab, \mval_k, \mval_v) &= (\mkont', \mathbf{snd}(\lfp({\mathcal M}(\mctx)))) \text{ if } (\mkont',\mctx) = \domarkstep(\mkont) \\
  \text{where } {\mathcal M}(\mctx_0)(C, \mktab) &= (C \cup C' \cup \set{\mctx_0}, \mktab \sqcup \mktab') \\
    I &= \setbuild{(\rewritectx(\mctx), \setbuild{\domarkstep(\mkont')}{\mkont' \in \mktab(\mctx)})}{\mctx \in C} \\
    C' &= \setbuild{\mctx}{(\_, S) \in I, (\_, \mctx) \in S} \\
    \mktab' &= \bigsqcup\setbuild{[\mctx \mapsto \setbuild{\mkont}{(\mctx, S) \in I, (\mkont, \_) \in S}]}{(\mctx,\_) \in I}
\\[2pt]
  \rewritectx(\mexpr,\menv,\mstore,(\mmarks, \mstore')) &= (\mexpr,\menv,\mstore,(\mmarks\sqcup_{!{}}[\mval_k \mapsto \mval_v], \mstore'))
\\[2pt]
  \domarkstep(\krt{\mctx}^\mmarks) &= (\krt{\rewritectx(\mctx)}^{\mmarks \sqcup_{!{}} [\mval_k \mapsto \mval_v]}, \mctx) \\
  \domarkstep(\kcons{\kwcm{\mmarks}^{\mmarks'}}{\mkont}) &= (\kcons{\kwcm{\mmarks \sqcup_{!{}} [\mval_k \mapsto \mval_v]}^{\mmarks' \sqcup_{!{}} [\mval_k \mapsto \mval_v]}}{\mkont}, \bot) \\
  \domarkstep(\mkont^\mmarks) &= (\kcons{\kwcm{[\mval_k \mapsto \mval_v]}^{\mmarks \sqcup_{!{}} [\mval_k \mapsto \mval_v]}}{\mkont}, \bot)
\end{align*}

Second, the $\domarkset$ metafunction builds a map of keys to lists, and builds a private store for the list constructions.
%
For this, we extend $\alloc$'s specification to allow more arguments---the private store, and a set of keys that need cons cells allocated for them---and produce a map from keys to pairs of addresses.
%
%% FIXME: need to crawl through qualified continuations for all extensions to build a list for each found key.
%% Has another implicit equality check
\newcommand{\markaux}{\mathit{marks}}
\begin{align*}
  \domarkset(\mstate\equiv\tpl{\mpoint,\mstore,\mkont,\mktab,\mmemo}) &= \markaux(\mkont,\emptyset) \\
  \markaux(\kmt^\bot, G) &= (\bot, \bot) \\
  \markaux(\krt{\mctx}^\mmarks, G) &= (\bot, \bot) \text{ if } \mctx \in G \\ %% FIXME: unsound
  \markaux(\krt{\mctx}^\mmarks, G) &= \bigsqcup\limits_{\mkont\in \mktab(\mctx)}{\markaux(\mkont,G \cup \set{\mctx})} \text{ otherwise} \\
  \markaux(\kcons{\kwcm{\mmarks}^{\mmarks'}}{\mkont}, G) &= (\mmarks''[\mval_k \mapsto \set{\cons{a}{d}} : \mval_k \in \dom(\mmarks), (a, d) = A(\mval_k)], \\
    &\phantom{= (}
      \mstore\sqcup[a \mapsto \mmarks(\mval_k) : \mval_k \in \dom(\mmarks), (a,\_) = A(\mval_k)] \\
    &\phantom{= (\mstore}\sqcup[d \mapsto \mmarks''(\mval_k) : \mval_k \in \dom(\mmarks), (\_,d) = A(\mval_k)]
   \\ \text{where } (\mmarks'', \mstore) &= \markaux(\mkont, G)
   \\ A &= \alloc(\mstate, \mstore, \dom(\mmarks))
\end{align*}
}

@@

The meta-functions mentioned in the fixed semantics deal with the addition of $\maddr$ to contexts.
%
At capture time, we strip the store and stored continuations closure in the $\rt$ continuation if there is one, and replace them with an address to make the context storeable.
%
We can't recursively map addresses to stores and continuation closures, so we flatten the context's closure into the current one.
%
This operation makes lookups more costly since we have to find closures bounded above by the current closure, rather than one exact closure.

@@

This theorem amounts to just the first bullet of \autoref{thm:concrete-tabular}.
%
The definition of ${\mathcal F}$ is similar in that only $\mktab$ and $\mmemo$ are widened, but states in $R$ now also contain $\mmkont$s and $\mmktab$s.
%
The definition of $\reify$ must change to account for the different target state space, the new context form and the continuation closure:
\\

>>>>

%%
This paper gives a common, simple methodology to derive CFA2, PDCFA and its garbage-collecting extension in a manner similar to AAM.
%
Our presentation is meant to be a tutorial to this related literature, with the additional promise of a novel analysis.
%
We don't presume a language of formal semantics such that our technique applies to all formalisms in such a language.
%
An overly formal approach is unlikely to capture all ``real language's'' formalizations.
%
The additional abstraction also detracts from the operational intuitions and simplicity of the methodology.
%%

%%
We present the definitions and proofs in a manner that should be well-positioned for mechanization.
%
The reader unfamiliar with or uninterested in correctness proofs should still be able to understand the technique and apply it with little difficulty.
%
This paper has been partially mechanized in the Coq proof assistant~\citep{Filliâtre97thecoq}, with proofs in the supplemental materials\footnote{\url{http://github.com/ianj/concrete-summaries} in \texttt{model.v}}.

@@

CFA2 is an online version of a technique called ``summarization'' from Sharir and Pnueli \citep[Chapter 7]{local:muchnick:jones:flow-analysis:1981}, which is synonymous with PDCFA's crucial ``$\epsilon$-closure graph.''
%
The online quality of the algorithm is what makes it suitable for higher-order languages, since CFA essentially interprets the program for each discovered state to build the model on-the-fly.
%
Summarization algorithms need not be restricted to languages with well-bracketed calls and returns.
%
The technique is adaptable for higher precision in the common case but to still handle difficult cases such as first-class control.
%
This was shown for the \rackett{call/cc} operator by~\citet{ianjohnson:Vardoulakis2011Pushdown}.
%
This work illuminated the fact that we can harness the enhanced technology of pushdown analyses in non-pushdown models of computation.
%
Doing this sacrifices call/return matching in the general case, but in practice the precision is much better than the alternative regular model that, \eg, 0-CFA would provide.
%%

@@

%%
The second piece of machinery serves two purposes:
%
(1) the state space contains all function call-sites to project out a control-flow graph, and %
(2) the worklist semantics of execution makes it possible to appeal to an incomplete memo table, so the saved stacks allow new results to flow to calls that used incomplete memo information.
%
Once the machine semantics is in this form, simple point-wise abstraction leads to the summarization algorithms that we see in the literature.

@@

\paragraph{Notation:} we make judicious use of the isomorphism between associations of tuples to spare the paper from excessive parentheses, even though our implementations are in a Lisp-descendant language.

@@

%%
The tail of a continuation within a function is some $\krt{\mctx}$, which instructs the semantics to memoize the result of the function evaluation at $\mctx$.
%
A context, $\mctx$, must contain enough information for memoization to be sound, \ie, any extra components not in $\mctx$ must be irrelevant to the evaluation of the function.
%
In this case, the context includes the function, environment, and store.
%
All the extra components at a context $\mctx$ get stored in a separate table, $\mktab$.
%
Updates to the memo table will propagate to all callers that depend on the result of $\mctx$.
%%

%%
Continuations count as ``extra'' in this semantics since functions cannot inspect the stack, so $\krt{\mctx}$ might be considered a special allocation strategy for continuations in the AAM viewpoint.
%
We separate the table of continuations, $\mktab$, from the store, $\mstore$, since $\rt$ continuations contain stores---this leads to a recursive data-structure that can cause the analysis to diverge.
%
The store is an important ingredient to maintaining enough precision to keep a pushdown abstraction.
%
The semantic spaces for the machine are modified as follows:

\begin{align*}
  \mstate \in \State &::= \tpl{\mexpr,\menv,\mstore,\mkont,\mktab,\mmemo} \\
  \mkont \in \Kont &::= \kmt \alt \krt{\mexpr,\menv,\mstore} \alt \kcons{\mkframe}{\mkont} \\
  \mktab \in \KTab &= \Expr \times \Env\times \Store \to \wp(\Kont) \\
  \mmemo \in \Memo &= \Expr \times \Env\times \Store \to \wp(\Value \times \Store)
\end{align*}

@@

\paragraph{The role of summaries:} notice the additional $\Memo$ component.
%
A ``summary edge'' in CFA2 or equivalently an $\epsilon$-edge in PDCFA is an edge from the source of a push edge to the target of the matching pop edge.
%
``Matching'' here means there is a path through machine reductions that don't change the stack, or through summary edges.
%
In our model, we only ``push'' when we call a function, and we only ``pop'' when we return with a value.
%
There is an analogy to something more operational: summary edges embody memoization.
%
Outside the pure world, functions don't simply return values --- they can also impact the outer environment.
%
In our setting of the CESK machine, we also memoize the state of the store.
%
Once we reach a function call for which we have witnessed a result, we can simply jump to that result (with the current continuation).
%
This is the second case of the function call rule in \autoref{fig:summary-semantics}
%%

\paragraph{A fly in the ointment} This semantics is only correct for fresh allocation strategies, which in turn make the reduction relation deterministic.
%
As it is, the CESK machine approximates the tabular machine, but not the other way around;
%
if we use a finite allocation strategy for example, then we don't explore some non-deterministic paths that lead to different results for a function call.
%
We thus explore all paths and share $\mktab$ and $\mmemo$ between all paths.

@@

{
\setlength{\abovedisplayskip}{0pt}
\setlength{\belowdisplayskip}{4pt}
\setlength{\abovedisplayshortskip}{0pt}
\setlength{\belowdisplayshortskip}{8pt}
\begin{align*}
  \mastate \in \sa{State} &::= \tpl{\mexpr,\menv,\mstore,\mkont} \\
  \System &= \wp(\sa{State}) \times \wp(\sa{State} \times \sa{State}) \times \KTab \times \Memo \\
  {\mathcal F} &: \sa{\State} \to \System \to \System
\end{align*}
\begin{align*}
  {\mathcal F}(\mastate_0)(\emptyset, R, \mktab, \mmemo) &= (\emptyset, R, \mktab, \mmemo) \\
  {\mathcal F}(\mastate_0)(F\uplus\set{\mastate}, R, \mktab, \mmemo) &= (F\cup F', R \cup R', \mktab', \mmemo') \\
  \text{where } I &= \setbuild{\mstate'}{\wn(\mastate,\mktab,\mmemo) \stepto \mstate'} \\
                \wn(\tpl{\mexpr,\menv,\mstore,\mkont},\mktab,\mmemo)
                   &= \tpl{\mexpr,\menv,\mstore,\mkont,\mktab,\mmemo}
\end{align*}
\begin{align*}
    R' &= \setbuild{(\mastate,\mastate')}{\wn(\mastate', \_, \_) \in I \wedge (\mastate,\mastate') \text{ not a call using } \mmemo} \\
    \mktab' &=  \bigsqcup\setbuild{\mktab'}{\wn(\_, \mktab', \_) \in I} \\
    \mmemo' &= \bigsqcup\setbuild{\mmemo'}{\wn(\_, \_, \mmemo') \in I} \\
    F' &= (\setbuild{\mastate'}{\wn(\mastate', \_, \_) \in I} \cup \set{\mastate_0})\setminus (\pi_1 R \cup \pi_2 R)
  \end{align*}
}

A $\System$ embodies the reductions seen disregarding the administrative pop steps, $R$, the frontier set of states yet to analyze, $F$, the continuation table $\mktab$ and the memo table $\mmemo$.
%
The set of reductions does not include uses of the memo table since those do not correspond to true steps that the original machine takes.
%
We define $\pi_1$ and $\pi_2$ as the point-wise lifting of the left and right projections out of a pair, respectively.
%
The least fixed-point of ${\mathcal F}(\tpl{\mexpr,\bot,\bot,\kmt})$ is the analysis of closed expression $\mexpr$.
%
Note the ordering on systems is point-wise, so the bottom element is $(\emptyset,\emptyset,\bot,\bot)$, which we'll call $\mathit{sys}_0$.
%
Systematic techniques for a performant implementation can be found in \citet{ianjohnson:oaam:icfp2013}.
%

\paragraph{Technical aside} There are minor differences with PDCFA's algorithm due to our choice to process one state at a time. If we process the entire frontier in one step, then we need to have additional tables of ``changes to $\mktab$ and $\mmemo$'' that need to be matched up in the following step. Here ``matched up'' means if both tables have the same context in their domain, we must add the state that has the memoized result(s) with the saved continuation(s). If we don't, then some memoized results may not flow to newly reached calling contexts. This corresponds with PDCFA's worksets of edges.
%%

% Note: Only if widened store.{
%
%The memo table need only map to values, since the global store will always subsume the store that would otherwise be mapped.
%
%All the states in the frontier are stepped with the current store, after which the next store is the least upper bound of all the resulting stores.
%
%The next frontier contains only states resulting from stepping the previous frontier that we haven't seen at this next store.}
%

@@

\paragraph{Example:} Let's consider a monovariant allocation strategy to run the following example\footnote{The example is first-order for brevity, not technical difficulty.}.
Note that monovariance means the address for each binding is the variable name itself, so $\menv$ is always the identity function. We therefore omit it.

@@

%%
We should expect that a pushdown analysis would predict this evaluates to true, and there are no loops in the program.
%
0-CFA~\citep{dvanhorn:Shivers:1991:CFA} claims there is a loop from the second call of \texttt{id} to the first, and thus predicts this program evaluates to true or false.
%
PDCFA claims there is no loop and that the result is true.
%
The example evaluation is in \autoref{fig:ex-eval}.
%
We maintain enough context to distinguish the return points of \texttt{id} to not rebind \texttt{y} to 1.
%
When determining control flow through the expression, we consult $\mktab$ to continue past function returns, so there is no confusion about a back edge from the second call to \texttt{id}.
%

@@

Garbage collection ($\Gamma$) restricts the heap to just the reachable addresses, so a semantics with GC introduces the following reduction rule:
\begin{center}
  \begin{minipage}{0.3\linewidth}
    \begin{align*}
      \mstate \stepto \Gamma(\mstate)
    \end{align*}
  \end{minipage}
  \begin{minipage}{0.65\linewidth}
    \begin{align*}
      \Gamma(\tpl{\mexpr,\menv,\mstore,\mkont, \mktab, \mmemo}) &= \tpl{\mexpr,\menv,\mstore|_{\reaches(\mathit{root}, \mstore)}, \mkont, \mktab, \mmemo} \\
      \text{where } \mathit{root} &= \touchesk{\mkont}(\mktab) \cup
      \mathit{range}(\menv|_{\mathit{fv}(\mexpr)})
    \end{align*}
  \end{minipage}
\end{center}
%
The ``reaches'' meta-function ($\reaches$) is a transitive closure of ``touches'' as it weaves through the store, starting from some root set of addresses:

@@

This semantics of course enjoys the soundness properties that GC does not introduce dangling pointers, equal length traces are equal up to GC (in the concrete), and abstractions are sound with respect to collected states.
%
The proofs are all similar to the original abstract GC paper's.
%
Technically since this semantics inspects the touched addresses of the continuation, the set of touched addresses should be considered part of the ``context'' in the $\mathbf{rt}$ continuations.
%
This extra context can be overlooked, but it leads to larger live address sets.
%
More live addresses means a less effective GC, but the introspective pushdown analysis chose this trade-off for a possibly smaller state space.
%

@@

\begin{enumerate}
\item{stack allocation for some bindings in an additional $\mframe \in \Store$}
\item{strong updates on stack frames for looked-up addresses}
\end{enumerate}
There are two orthogonal features of the semantics, embodied by our semantics' two meta-functions ($\bind$ and $\lookup$, respectively in \autoref{fig:frame-semantics}).

@@

%%
The stack frame introduces a refinement over the original CESK machine, so we first prove it correct with respect to the concrete semantics, in a similar fashion to PDCFA.
%
\begin{theorem}[Correctness of stack frame refinement]\label{thm:refinement}
  The CES$\xi$K machine has a stuttering bisimulation refinement with the CESK machine with fresh allocation.
\end{theorem}
This follows from the invariant that if a stack frame maps $\maddr$ to $\mval$, then the store maps $\maddr$ to $\mval$, meaning that the special lookup function is equivalent to lookup directly in the store (exploits immutability).
%
We stutter past the stack frame replacement steps.
%
The refinement simply removes the local stack frame from the state and $\mathbf{push}$ frames from the stack.
%
%That is, if we always allocated bindings in the heap.
%
%The additional escape information that allows us to elide allocation in the heap introduces a second level of reasoning, where the burden of proof falls on the escape analysis.
%
%The escape analysis adds information to use in the proof, that if $\mvar$ never escapes, then every reference to allocations of $\mvar$'s binding will always be able to use $\mframe$ for the lookup.
%
%With this proposition, we can prove that the CES$\xi$K machine that always allocates in the store has a bisimulation with the refined machine that exploits escape analysis to avoid unnecessary changes to the store.
%

@@

The two meta-functions, $\domark$ and $\domarkset$ are what do the heavy-lifting of adding marks to a continuation and reifying the mark set into a map of keys to lists.
%
Adding a mark to a continuation has an implicit equality check; if $\mval_k = \mval_k'$ concretely, then {\tt wcm} should do a strong update on $\mmarks$, but otherwise it should join since they could be two different values.
%
Concrete equality analysis is outside the scope of this paper, but doable (\eg, with abstract counting~\citep{dvanhorn:Might:2006:GammaCFA}), so we elide the details and call this strong-update-or-join operation $\sqcup_{!{}}$.
%
Marking a return continuation means we need to crawl through the table, which again can be circular, so we compute the least fixed point of ${\mathcal M}$ to get an extended table holding the marked continuations.
%
\newcommand{\rewritectx}{\mathit{rw}}
\newcommand{\domarkstep}{\mathit{mark}_0}
\begin{align*}
  \domark(\mkont, \mktab, \mval_k, \mval_v) &= (\mkont', \mktab) \text{ if } (\mkont',\bot) = \domarkstep(\mkont) \\
  \domark(\mkont, \mktab, \mval_k, \mval_v) &= (\mkont', \mathbf{snd}(\lfp({\mathcal M}(\mctx)))) \text{ if } (\mkont',\mctx) = \domarkstep(\mkont) \\
  \text{where } {\mathcal M}(\mctx_0)(C, \mktab) &= (C \cup C' \cup \set{\mctx_0}, \mktab \sqcup \mktab') \\
    I &= \setbuild{(\rewritectx(\mctx), \setbuild{\domarkstep(\mkont')}{\mkont' \in \mktab(\mctx)})}{\mctx \in C} \\
    C' &= \setbuild{\mctx}{(\_, S) \in I, (\_, \mctx) \in S} \\
    \mktab' &= \bigsqcup\setbuild{[\mctx \mapsto \setbuild{\mkont}{(\mctx, S) \in I, (\mkont, \_) \in S}]}{(\mctx,\_) \in I}
\\[2pt]
  \rewritectx(\mexpr,\menv,\mstore,(\mmarks, \mstore')) &= (\mexpr,\menv,\mstore,(\mmarks\sqcup_{!{}}[\mval_k \mapsto \mval_v], \mstore'))
\\[2pt]
  \domarkstep(\krt{\mctx}^\mmarks) &= (\krt{\rewritectx(\mctx)}^{\mmarks \sqcup_{!{}} [\mval_k \mapsto \mval_v]}, \mctx) \\
  \domarkstep(\kcons{\kwcm{\mmarks}^{\mmarks'}}{\mkont}) &= (\kcons{\kwcm{\mmarks \sqcup_{!{}} [\mval_k \mapsto \mval_v]}^{\mmarks' \sqcup_{!{}} [\mval_k \mapsto \mval_v]}}{\mkont}, \bot) \\
  \domarkstep(\mkont^\mmarks) &= (\kcons{\kwcm{[\mval_k \mapsto \mval_v]}^{\mmarks \sqcup_{!{}} [\mval_k \mapsto \mval_v]}}{\mkont}, \bot)
\end{align*}

Second, the $\domarkset$ meta-function builds a map of keys to lists, and builds a private store for the list constructions.
%
For this, we extend $\alloc$'s specification to allow more arguments---the private store, and a set of keys that need cons cells allocated for them---and produce a map from keys to pairs of addresses.
%
%% FIXME: need to crawl through qualified continuations for all extensions to build a list for each found key.
%% Has another implicit equality check
\newcommand{\markaux}{\mathit{marks}}
\begin{align*}
  \domarkset(\mstate\equiv\tpl{\mpoint,\mstore,\mkont,\mktab,\mmemo}) &= \markaux(\mkont,\emptyset) \\
  \markaux(\kmt^\bot, G) &= (\bot, \bot) \\
  \markaux(\krt{\mctx}^\mmarks, G) &= (\bot, \bot) \text{ if } \mctx \in G \\ %% FIXME: unsound
  \markaux(\krt{\mctx}^\mmarks, G) &= \bigsqcup\limits_{\mkont\in \mktab(\mctx)}{\markaux(\mkont,G \cup \set{\mctx})} \text{ otherwise} \\
  \markaux(\kcons{\kwcm{\mmarks}^{\mmarks'}}{\mkont}, G) &= (\mmarks''[\mval_k \mapsto \set{\cons{a}{d}} : \mval_k \in \dom(\mmarks), (a, d) = A(\mval_k)], \\
    &\phantom{= (}
      \mstore\sqcup[a \mapsto \mmarks(\mval_k) : \mval_k \in \dom(\mmarks), (a,\_) = A(\mval_k)] \\
    &\phantom{= (\mstore}\sqcup[d \mapsto \mmarks''(\mval_k) : \mval_k \in \dom(\mmarks), (\_,d) = A(\mval_k)]
   \\ \text{where } (\mmarks'', \mstore) &= \markaux(\mkont, G)
   \\ A &= \alloc(\mstate, \mstore, \dom(\mmarks))
\end{align*}
}

@@

The meta-functions mentioned in the fixed semantics deal with the addition of $\maddr$ to contexts.
%
At capture time, we strip the store and stored continuations closure in the $\rt$ continuation if there is one, and replace them with an address to make the context storeable.
%
We can't recursively map addresses to stores and continuation closures, so we flatten the context's closure into the current one.
%
This operation makes look-ups more costly since we have to find closures bounded above by the current closure, rather than one exact closure.

@@

\paragraph{Notice} implementers may skip to the next bolded subsection.
%%

%%
This theorem amounts to just the first bullet of \autoref{thm:concrete-tabular}.
%
The definition of ${\mathcal F}$ is similar in that only $\mktab$ and $\mmemo$ are widened, but states in $R$ now also contain $\mmkont$s and $\mmktab$s.
%
The definition of $\reify$ must change to account for the different target state space, the new context form and the continuation closure:
\\
